---
title: "Activity 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
```
# Day 1

## Task 3: Load the data and

```{r}
hfi <- read_csv("https://www.openintro.org/data/csv/hfi.csv");
glimpse(hfi)
head(hfi)
```
### Dataset

There are 123 columns/variables and 1458 observations.

Each row in the dataset gives the freedom score, rank, quantile and other scores for a single country for a year.

### Filtering for only 2016

```{r}
hfi_2016 <- hfi %>% filter(year == 2016)
head(hfi_2016)
```
### Plot

```{r}
hfi_2016 %>% ggplot(aes(pf_expression_control, pf_score)) +
  geom_point() +
  labs(
    title = "Personal Freedom (score) vs Political pressures and controls \non media content for year 2016",
    x = "Political pressures and controls on media content",
    y = "Personal Freedom (score)"
  )+
  theme_bw()
```

The relationship looks linear with a positive slope. A linear model seems to be a good choice here.


## Task 4: Sum of squared residuals

The manual process gave me a minimum sum of squares of 106.83 for line pf_score = 4.1750 + 0.5302 * pf_expression_control.

## Task 5: The linear model

```{r}
m1 <- lm(pf_score ~ pf_expression_control, data = hfi_2016)
tidy(m1)
```

Here the y-intercept is: 4.2838153 and the slope is: 0.5418452.

# Day 2

## Task 2: Overall model fit

### Correlation

```{r}
cor(hfi_2016$pf_score, hfi_2016$pf_expression_control)
```
This means that pf_score and pf_expression_control have high positive correlation. That means if pf_expression_control increases pf_score also increases and we can use pf_expression_control to predict pf_score.

### R^2^ of the model

```{r}
glance(m1)
```

The R^2^ of the model is 0.7141342.

This means that 71.41% of the variability in pf_score is explained away by this model.

### Predicting Human freedom score based on Political pressures and controls on media content

```{r}
m2 <- lm(hf_score ~ pf_expression_control, data = hfi_2016)
tidy(m2)
```

When no Political Pressure and Control is exerted on the media content, the mean Human freedom score estimate is 5.053396. 

For unit increase in Political pressure and control the increase in mean Human freedom score is estimated to be 0.368432.

Political pressure and control on the media content and the Human freedom score have a positive relationship.

## Task 3: Prediction and prediction errors

### Plot

```{r}
hfi_2016 %>% ggplot(aes(pf_expression_control, pf_score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Personal Freedom (score) vs Political pressures and controls \non media content for year 2016",
    x = "Political pressures and controls on media content",
    y = "Personal Freedom (score)"
  )+
  theme_bw()
```

From just the plot, to get Personal Freedom score for 3 rating for pf_expression_control, we can follow the vertical line at 3 on the x-axis until we get to the lm line. It gives approximately 6.3 for pf_score.

This is actually an overestimate becuase most of the points in the data is below the line.

## Task 4: Model diagnostics

### Residual Plot

```{r}
m1_aug <- augment(m1)

ggplot(data = m1_aug, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted values") +
  ylab("Residuals")
```

The residuals have a random spread from left to right with most points within 1 to -1. The residuals are also closer to 0. This means that pf_expression_control is a good predictor of pf_score.


### Residuals Histogram

```{r}
ggplot(data = m1_aug, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")
```

The histogram of the residuals is unimodal and skewed left. This raises questions about the normality of the data.

Looking at the residual vs fitted plot it appears that the constant variability assumption is met since there are no apparent pattern visible.

## Challenge: More practice

I choose _pf_expression_ which is a score of Freedom of expression to predict _pf_score_. I chose _pf_expression_ because freedom of expression is important for personal freedom.

### Plot
```{r}
hfi_2016 %>% ggplot(aes(pf_expression, pf_score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Personal Freedom (score) vs Freedom of expression for year 2016",
    x = "Freedom of expression",
    y = "Personal Freedom (score)"
  )+
  theme_bw()
```

At a glance the relationship seems linear though fitting a quadratic model may give better R^2^.

### Model

```{r}
m3 <- lm(pf_score~pf_expression, hfi_2016)
summary(m3)
```

The model with pf_expression is significant so there is a linear relationship between pf_expression and pf_score.

### R^2^ comparision

The model with pf_expression_control and pf_score has a R^2^ of 0.7141342 which is arguably same as that of this model (0.7139). So this model is not better at predicting pf_score than the previous one.

### Checking Regression assumptions

```{r}
m3_aug <- augment(m3)

# Residuals vs Fitted plot

ggplot(data = m3_aug, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted values") +
  ylab("Residuals")

# Histogram of Residuals
ggplot(data = m3_aug, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")
```

We can argue that there's a pattern to the residual vs fitted plot. But the pattern is not apparent and the points do cluster near to 0 along y-axis.

The histogram is skewed left. So both the normality assumption and the constant variance assumption are under question and requires further investigation.

```{r}
shapiro.test(hfi_2016$pf_expression)
shapiro.test(hfi_2016$pf_score)
```

Shapiro-Wilk normality test shows that the pf_expression data is not normal. The same is true for pf_score also.

### Another Set of Variables

I'm choosing to predict Economic freedom score (`ef_score`) based on the Business regulations - Extra payments/bribes/favoritism (`ef_regulation_business_bribes`). This is an interesting case because I wonder if greasing your way to better business deal means more economic freedom.

### Plot and Model fit

```{r}
# Plot
hfi_2016 %>% ggplot(aes(ef_regulation_business_bribes, ef_score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Personal Freedom (score) vs Business regulations - \nExtra payments/bribes/favoritism for year 2016",
    x = "Business regulations",
    y = "Personal Freedom (score)"
  )+
  theme_bw()

# Model
m4 <- lm(ef_score~ef_regulation_business_bribes, hfi_2016)
summary(m4)
```
Though the relationship is linear and the model is significant, the R^2^ is low (0.395). This means Economic freedom depends on many other things and just having a huge freedom to Extra payments/bribes/favoritism doesn't actually give greater economic freedom.

Though it is surprising that there's a positive relationship between the two. I expected it to be a negative. Maybe greater value of ef_regulation_business_bribes means that people are forced to make illegal deals less therefore have greater freedom from extra payments, bribes, favoritism etc.


